Manuscript

### 2018-11-15 manuscript.Rmd
rm old data files, make new 18Q4 Rds

@@

WilcoxonByDrugAllGenes <- function(melt_file, grid_file, data_name) {
  druglist <- as.character(unique(melt_file$Drug))
  sig_df <- NULL
  for(i in 1:length(druglist)) {
    print(paste0("Current drug in process: ", druglist[i], " (drug ", i, " of ", length(druglist), ")."))
    flush.console()
    melted <- filter(melt_file, Drug == druglist[i])
    use_data <- merge(grid_file, melted, by = "accession_id", all.x = TRUE)
    use_data <- merge(use_data, ccl_converter, by = "accession_id", all.x = TRUE)
    use_data <- merge(use_data, maf_df, by = c("Hugo_Symbol", "CCLE_Name", "Broad_ID"), all.x = TRUE)
    use_data$Mutation_Status_Nonsilent <- ifelse(is.na(use_data$Mutation_Status_Nonsilent), "Other", use_data$Mutation_Status_Nonsilent)
    use_data$Hugo_Symbol <- factor(use_data$Hugo_Symbol)
    use_data$AUC <- as.numeric(use_data$AUC)
    use_data <- filter(use_data, !is.na(use_data$AUC))
    use_data$Drug_Gene <- paste0(use_data$Drug, "_", use_data$Hugo_Symbol)
    use_data_ptmuts <- filter(use_data, Reference_Allele_Length == 1 | Reference_Allele_Length == 0 | is.na(Reference_Allele_Length))
    use_data_ptmuts$Keep_PtMut_Tests <- ifelse(use_data_ptmuts$Reference_Allele_Length == 1 | is.na(use_data_ptmuts$Reference_Allele_Length), TRUE, ifelse(use_data_ptmuts$Tumor_Seq_Allele1_Length == 1, TRUE, FALSE))
    use_data_ptmuts <- filter(use_data_ptmuts, Keep_PtMut_Tests == TRUE)

@@

## Pearson correlation by lineage
```{r}
PearsonByLineage <- function(lin, dataset, variable) {
  use_data <- filter(dataset, group_general_lineage_name == lin)
  if(variable == "GE") {
    use_data <- filter(use_data, !is.na(RPKM))
    use_data_summ <- use_data %>% count(Drug_Gene)
    use_data_summ <- filter(use_data_summ, n >= 3) # cor.test requires at least 3 non-NA data points
    use_data_filt <- filter(use_data, Drug_Gene %in% as.character(unique(use_data_summ$Drug_Gene)))
    if(nrow(use_data_filt) > 0) {
      pearson_res <- use_data_filt %>% group_by(Drug_Gene) %>% summarize(Pearson_Corr_GE = cor.test(y = AUC, x = RPKM, method = "pearson", use = "complete.obs")$estimate, Pearson_Pval_GE = cor.test(y = AUC, x = RPKM, method = "pearson", use = "complete.obs")$p.value)
      pearson_res <- merge(g2p_druggene_p_tally, pearson_res, all.y = TRUE)
      pearson_res$group_general_lineage_name <- lin
      pearson_res$Pearson_GE_Abs <- abs(pearson_res$Pearson_Corr_GE)
      pearson_res$Pearson_GE_Sign <- ifelse(sign(pearson_res$Pearson_Corr_GE) == 1, "+",
                                              ifelse(sign(pearson_res$Pearson_Corr_GE) == -1, "-", "0"))
      return(pearson_res)
    }
  }
  else if(variable == "CN") {
    use_data <- filter(use_data, !is.na(Copy_Number))
    use_data_summ <- use_data %>% count(Drug_Gene)
    use_data_summ <- filter(use_data_summ, n >= 3) # cor.test requires at least 3 non-NA data points
    use_data_filt <- filter(use_data, Drug_Gene %in% as.character(unique(use_data_summ$Drug_Gene)))
    if(nrow(use_data_filt) > 0) {
      pearson_res <- use_data_filt %>% group_by(Drug_Gene) %>% summarize(Pearson_Corr_CN = cor.test(y = AUC, x = Copy_Number, method = "pearson", use = "complete.obs")$estimate, Pearson_Pval_CN = cor.test(y = AUC, x = Copy_Number, method = "pearson", use = "complete.obs")$p.value)
      pearson_res <- merge(g2p_druggene_p_tally, pearson_res, all.y = TRUE)
      pearson_res$group_general_lineage_name <- lin
      pearson_res$Pearson_CN_Abs <- abs(pearson_res$Pearson_Corr_CN)
      pearson_res$Pearson_CN_Sign <- ifelse(sign(pearson_res$Pearson_Corr_CN) == 1, "+",
                                              ifelse(sign(pearson_res$Pearson_Corr_CN) == -1, "-", "0"))
      return(pearson_res)
    }
  }
}
```

@@

This comprehensive cancer cell line information curated by Daniel Charytonowicz.
```{r}
ccl_converter <- read.delim("./data_munging/cell_line_database_v1_20180911.tsv", row.names = 1, sep = "\t", header = TRUE)
lineage_converter <- unique(ccl_converter[, c("grouped_general_doid", "group_general_lineage_name")])
```

@@

```{r}
ccl_info <- read.delim("./data_munging/DepMap-2018q3-celllines.csv", sep = ",", header = TRUE, na.strings = c("", NA))
ccl_info$Primary.Disease <- gsub("\\\\", "", ccl_info$Primary.Disease)
ccl_info$Primary.Disease <- gsub("Ewings", "Ewing's", ccl_info$Primary.Disease)

# From figshare
crispr_meta <- read.delim("./data_munging/sample_info_18Q3_crispr.csv", sep = ",", header = TRUE, na.strings = c("", NA))
colnames(crispr_meta)[7] <- "CCLE_Name"
```

@@

For mutation calling, get paired gene name and cell line fields in a data frame. For this analysis, we don't care about how many mutations there are per gene or what type of mutations there are, so I didn't save more information. I took unique gene-cell line combinations since we only cared about mutation presence/absence. Add a `Mutation_Status` column denoting all entries in MAF files as mutations present in the associated cell lines.
```{r}
maf_raw <- read.delim("./data_munging/CCLE_DepMap_18q3_maf_20180718.txt.gz", header = TRUE, sep = "\t")
# Filter for cell lines in CRISPR screen
maf_raw <- filter(maf_raw, Broad_ID %in% unique(crispr_meta$Broad_ID))
colnames(maf_raw)[colnames(maf_raw) == "Tumor_Sample_Barcode"] <- "CCLE_Name"

# Select columns
maf_df <- subset(maf_raw, select = c("Hugo_Symbol", "CCLE_Name", "Broad_ID", "Variant_Classification", "Reference_Allele", "Tumor_Seq_Allele1"))
maf_df$Reference_Allele_Length <- ifelse(maf_df$Reference_Allele == "-", 0, nchar(as.character(maf_df$Reference_Allele)))
maf_df$Tumor_Seq_Allele1_Length <- nchar(as.character(maf_df$Tumor_Seq_Allele1))

# Add Mutation_Status column
maf_df$Mutation_Status_Nonsilent <- ifelse(maf_df$Variant_Classification == "Silent", "Other", "Mutant")
maf_df <- unique(subset(maf_df, select = c("Hugo_Symbol", "CCLE_Name", "Broad_ID", "Mutation_Status_Nonsilent", "Reference_Allele_Length", "Tumor_Seq_Allele1_Length")))
```

@@

```{r, eval=FALSE}
cn <- read.delim("./data_munging/public_18Q3_gene_cn_v2.csv.gz", sep = ",", check.names = FALSE, header = TRUE)

# Convert log2 ratios (log2(CN/2)) to CN
cn[2:ncol(cn)] <- lapply(cn[2:ncol(cn)], function(x) 2 * (2 ^ x))

# Remove Entrez gene IDs from colnames
colnames(cn) <- gsub(" .*", "", colnames(cn))
colnames(cn)[1] <- "Broad_ID"

# Melt
cn_melt <- melt(data = cn, id.vars = "Broad_ID", measure.vars = colnames(cn[2:ncol(cn)]), variable.name = "Hugo_Symbol", value.name = "Copy_Number")

saveRDS(cn_melt, "./data_munging/rds/cn_melt_18Q3.rds", compress = "xz")
```

```{r, eval=FALSE}
cn_melt <- readRDS("./data_munging/rds/cn_melt_18Q3.rds")
```

@@

Gene expression data (Reads Per Kilobase of transcript, per Million mapped reads, RPKM).
```{r, eval=FALSE}
ge <- read.delim("./../crispr_lineages_giant_files/CCLE_DepMap_18q3_RNAseq_RPKM_20180718.gct.gz", skip = 2, header = TRUE, sep = "\t", check.names = FALSE)

# Edit columns
ge$Name <- NULL
colnames(ge)[1] <- "Hugo_Symbol"

# Melt
ge_melt <- melt(data = ge, id.vars = "Hugo_Symbol", measure.vars = colnames(ge[2:ncol(ge)]), value.name = "RPKM")
## Split variable column
ge_melt <- with(ge_melt, cbind(Hugo_Symbol, colsplit(variable, pattern = " ", names = c("CCLE_Name", "Broad_ID")), RPKM))
## Remove parentheses around Broad IDs
ge_melt$Broad_ID <- gsub("\\(|\\)", "", ge_melt$Broad_ID)

saveRDS(ge_melt, "./../crispr_lineages_giant_files/ge_melt_18Q3.rds", compress = "xz")
```

```{r, eval=FALSE}
ge_melt <- readRDS("./../crispr_lineages_giant_files/ge_melt_18Q3.rds")
```

@@

```{r, eval=FALSE}
dep <- read.delim("./data_munging/gene_dependency_18Q3.csv.gz", sep = ",", header  = TRUE, check.names = FALSE)
# Remove Entrez gene IDs from colnames
colnames(dep) <- gsub(" .*", "", colnames(dep))
```

@@

The latest CRISPR CERES score data (18Q3, August 2018) was pulled from the [DepMap Data Portal](https://depmap.org/portal/download/) (Broad Institute Cancer Dependency Map 2018, Meyers et al. 2017).
```{r, eval=TRUE}
crispr <- read.delim("./data_munging/gene_effect_18Q3.csv.gz", sep = ",", header  = TRUE, check.names = FALSE)
# Remove Entrez gene IDs from colnames
colnames(crispr) <- gsub(" .*", "", colnames(crispr))
```

Merge annotation data:
```{r, eval=FALSE}
# Melt CRISPR dataset for merging
crispr_melt <- melt(crispr, id.vars = "Broad_ID", measure.vars = colnames(crispr)[2:ncol(crispr)], variable.name = "Hugo_Symbol", value.name = "Score")
# Melt dependency probabilities dataset for merging
dep_melt <- melt(dep, id.vars = "Broad_ID", measure.vars = colnames(dep)[2:ncol(dep)], variable.name = "Hugo_Symbol", value.name = "Dep_Prob")
# Merge dependency probabilities
crispr_melt <- merge(crispr_melt, dep_melt, by = c("Broad_ID", "Hugo_Symbol"), all.x = TRUE)
# Merge cell line metadata
crispr_melt <- merge(crispr_melt, ccl_info, by = "Broad_ID", all.x = TRUE)
crispr_melt <- merge(crispr_melt, crispr_meta, by = c("CCLE_Name", "Broad_ID"), all.x = TRUE)
# Merge mutation annotations
crispr_muts <- merge(crispr_melt, maf_df, by = c("Hugo_Symbol", "CCLE_Name", "Broad_ID"), all.x = TRUE)
crispr_muts$Hugo_Symbol <- factor(crispr_muts$Hugo_Symbol)
crispr_muts <- crispr_muts %>% mutate(Mutation_Status_Nonsilent = if_else(is.na(Mutation_Status_Nonsilent), "Other", Mutation_Status_Nonsilent))
# Summarize number of mutant and Other cell lines
crispr_muts_summ <- crispr_muts %>% group_by(Hugo_Symbol) %>%
  summarize(N_Nonsilent_Other = sum(Mutation_Status_Nonsilent == "Other"),
            N_Nonsilent_Mutant = sum(Mutation_Status_Nonsilent == "Mutant"))
# Merge test results back into full dataset, which restores information lost in the summarization
crispr_data <- merge(crispr_muts_summ, crispr_muts, by = "Hugo_Symbol")
# Add Color columns
crispr_data$Color_Nonsilent <- ifelse(crispr_data$Mutation_Status_Nonsilent == "Other", "thistle", "darkorchid")
crispr_data$Color_Nonsilent <- factor(crispr_data$Color_Nonsilent)
# Cell line lineages
crispr_data <- merge(crispr_data, ccl_converter, by = c("CCLE_Name", "Broad_ID"), all.x = TRUE)
levels(crispr_data$lineage_name) <- sort(levels(crispr_data$lineage_name), decreasing = TRUE)
# Copy number
crispr_data <- merge(crispr_data, cn_melt, by = c("Hugo_Symbol", "Broad_ID"), all.x = TRUE)
# Gene expression (RPKM)
ge_filt <- filter(ge_melt, Hugo_Symbol %in% unique(crispr_data$Hugo_Symbol) & Broad_ID %in% unique(crispr_data$Broad_ID))
crispr_data <- merge(crispr_data, ge_filt, by = c("Hugo_Symbol", "Broad_ID", "CCLE_Name"), all.x = TRUE)
crispr_data$RPKM_log2 <- log2(crispr_data$RPKM + 0.0001)
saveRDS(crispr_data, "./../crispr_lineages_giant_files/crispr_data_18Q3.rds", compress = "xz")
```

```{r}
crispr_data <- readRDS("./../crispr_lineages_giant_files/crispr_data_18Q3.rds")

@@

```{r, eval=FALSE}
# Drug metadata
drug_meta <- read.delim("./data_munging/drug_id_drug_name_table_10_12_2018.tsv", sep = "\t", header = FALSE)
colnames(drug_meta) <- c("CID", "Drug")
drug_meta$Drug <- tolower(drug_meta$Drug)
maf_df_g2p <- filter(maf_df, Hugo_Symbol %in% g2p_genes)
ge_filt <- filter(ge_melt, Hugo_Symbol %in% g2p_genes)
cn_filt <- filter(cn_melt, Hugo_Symbol %in% g2p_genes)

# CCLE
ccle <- read.delim("./data_munging/data_drug_auc_ccle_10_12_2018.csv", sep = "\t", header  = TRUE, row.names = 1, check.names = FALSE)
ccle <- merge(drug_meta, ccle, by = "CID")
ccle_melt <- melt(ccle, id.vars = c("CID", "Drug"), measure.vars = colnames(ccle)[3:ncol(ccle)], variable.name = "accession_id", value.name = "AUC")
ccle_grid <- expand.grid("accession_id" = unique(ccle_melt$accession_id), "Hugo_Symbol" = g2p_genes)
ccle_data <- merge(ccle_grid, ccle_melt, by = "accession_id", all.x = TRUE)
ccle_data <- merge(ccle_data, ccl_converter, by = "accession_id", all.x = TRUE)
ccle_data <- merge(ccle_data, maf_df_g2p, by = c("Hugo_Symbol", "CCLE_Name", "Broad_ID"), all.x = TRUE)
ccle_data$Mutation_Status_Nonsilent <- ifelse(is.na(ccle_data$Mutation_Status_Nonsilent), "Other", ccle_data$Mutation_Status_Nonsilent)
ccle_data <- merge(ccle_data, ge_filt, by = c("Hugo_Symbol", "CCLE_Name", "Broad_ID"), all.x = TRUE)
ccle_data <- merge(ccle_data, cn_filt, by = c("Hugo_Symbol", "Broad_ID"), all.x = TRUE)
ccle_data$Hugo_Symbol <- factor(ccle_data$Hugo_Symbol)
ccle_data$AUC <- as.numeric(ccle_data$AUC)
ccle_data$Color_Nonsilent <- ifelse(ccle_data$Mutation_Status_Nonsilent == "Other", "palegreen3", "springgreen4")
ccle_data$Color_Nonsilent <- factor(ccle_data$Color_Nonsilent)
ccle_data <- filter(ccle_data, !is.na(ccle_data$AUC))
ccle_data$Drug_Gene <- paste0(ccle_data$Drug, "_", ccle_data$Hugo_Symbol)
saveRDS(ccle_data, "./data_munging/rds/ccle_data_18Q3_g2pgenes.rds", compress = "xz")

>>>>

Edited `.Renviron` to avoid "Error: memory exhausted (limit reached?)" when melting the GE dataset, as per the suggestion from [StackOverflow](https://stackoverflow.com/questions/51295402/r-on-macos-error-vector-memory-exhausted-limit-reached)
```
cd ~
touch .Renviron
open .Renviron

# Added following line to .Renviron:
R_MAX_VSIZE=100Gb
```

@@

WilcoxonByDrugAllGenes <- function(melt_file, grid_file, data_name) {
  druglist <- as.character(unique(melt_file$Drug))
  sig_df <- NULL
  for(i in 1:length(druglist)) {
    print(paste0("Current drug in process: ", druglist[i], " (drug ", i, " of ", length(druglist), ")."))
    flush.console()
    melted <- filter(melt_file, Drug == druglist[i])
    use_data <- merge(grid_file, melted, by = "accession_id", all.x = TRUE)
    use_data <- merge(use_data, ccl_converter, by = "accession_id", all.x = TRUE)
    use_data <- merge(use_data, maf_df, by = c("Hugo_Symbol", "CCLE_Name", "DepMap_ID"), all.x = TRUE)
    use_data$Mutation_Status_Nonsilent <- ifelse(is.na(use_data$Mutation_Status_Nonsilent), "Wildtype", use_data$Mutation_Status_Nonsilent)
    use_data$Hugo_Symbol <- factor(use_data$Hugo_Symbol)
    use_data$AUC <- as.numeric(use_data$AUC)
    use_data <- filter(use_data, !is.na(use_data$AUC))
    use_data$Drug_Gene <- paste0(use_data$Drug, "_", use_data$Hugo_Symbol)
    use_data_ptmuts <- filter(use_data, Reference_Allele_Length == 1 | Reference_Allele_Length == 0 | is.na(Reference_Allele_Length))
    use_data_ptmuts$Keep_PtMut_Tests <- ifelse(use_data_ptmuts$Reference_Allele_Length == 1 | is.na(use_data_ptmuts$Reference_Allele_Length), TRUE, ifelse(use_data_ptmuts$Tumor_Seq_Allele1_Length == 1, TRUE, FALSE))
    use_data_ptmuts <- filter(use_data_ptmuts, Keep_PtMut_Tests == TRUE)

@@

## Pearson correlation by lineage
```{r}
PearsonByLineage <- function(lin, dataset, variable) {
  use_data <- filter(dataset, group_general_lineage_name == lin)
  if(variable == "GE") {
    use_data <- filter(use_data, !is.na(RPKM))
    use_data_summ <- use_data %>% count(Drug_Gene)
    use_data_summ <- filter(use_data_summ, n >= 3) # cor.test requires at least 3 non-NA data points
    use_data_filt <- filter(use_data, Drug_Gene %in% as.character(unique(use_data_summ$Drug_Gene)))
    if(nrow(use_data_filt) > 0) {
      pearson_res <- use_data_filt %>% group_by(Drug_Gene) %>% summarize(Pearson_Corr_GE = cor.test(y = AUC, x = RPKM, method = "pearson", use = "complete.obs")$estimate, Pearson_Pval_GE = cor.test(y = AUC, x = RPKM, method = "pearson", use = "complete.obs")$p.value)
      pearson_res <- merge(g2p_druggene_p_tally, pearson_res, all.y = TRUE)
      pearson_res$group_general_lineage_name <- lin
      pearson_res$Pearson_GE_Abs <- abs(pearson_res$Pearson_Corr_GE)
      pearson_res$Pearson_GE_Sign <- ifelse(sign(pearson_res$Pearson_Corr_GE) == 1, "+",
                                              ifelse(sign(pearson_res$Pearson_Corr_GE) == -1, "-", "0"))
      return(pearson_res)
    }
  }
  else if(variable == "CN") {
    use_data <- filter(use_data, !is.na(CN))
    use_data_summ <- use_data %>% count(Drug_Gene)
    use_data_summ <- filter(use_data_summ, n >= 3) # cor.test requires at least 3 non-NA data points
    use_data_filt <- filter(use_data, Drug_Gene %in% as.character(unique(use_data_summ$Drug_Gene)))
    if(nrow(use_data_filt) > 0) {
      pearson_res <- use_data_filt %>% group_by(Drug_Gene) %>% summarize(Pearson_Corr_CN = cor.test(y = AUC, x = CN, method = "pearson", use = "complete.obs")$estimate, Pearson_Pval_CN = cor.test(y = AUC, x = CN, method = "pearson", use = "complete.obs")$p.value)
      pearson_res <- merge(g2p_druggene_p_tally, pearson_res, all.y = TRUE)
      pearson_res$group_general_lineage_name <- lin
      pearson_res$Pearson_CN_Abs <- abs(pearson_res$Pearson_Corr_CN)
      pearson_res$Pearson_CN_Sign <- ifelse(sign(pearson_res$Pearson_Corr_CN) == 1, "+",
                                              ifelse(sign(pearson_res$Pearson_Corr_CN) == -1, "-", "0"))
      return(pearson_res)
    }
  }
}
```

@@

This comprehensive cancer cell line information curated by Daniel Charytonowicz.
```{r}
ccl_converter <- read.delim("./data_munging/cell_line_database_v1_20180911.tsv", row.names = 1, sep = "\t", header = TRUE)
lineage_converter <- unique(ccl_converter[, c("grouped_general_doid", "group_general_lineage_name")])
colnames(ccl_converter)[colnames(ccl_converter) == "Broad_ID"] <- "DepMap_ID"
```

@@

```{r}
ccl_info <- read.delim("./data_munging/18Q4/DepMap-2018q4-celllines_20181115.csv", sep = ",", header = TRUE, na.strings = c("", NA))
ccl_info$Primary.Disease <- gsub("\\\\", "", ccl_info$Primary.Disease)
ccl_info$Primary.Disease <- gsub("Ewings", "Ewing's", ccl_info$Primary.Disease)

# From figshare
crispr_meta <- read.delim("./data_munging/18Q4/sample_info_18Q4_crispr_20181115.csv", sep = ",", header = TRUE, na.strings = c("", NA))
colnames(crispr_meta)[7] <- "CCLE_Name"
```

@@

For mutation calling, get paired gene name and cell line fields in a data frame. For this analysis, we don't care about how many mutations there are per gene or what type of mutations there are, so I didn't save more information. I took unique gene-cell line combinations since we only cared about mutation presence/absence. Add a `Mutation_Status` column denoting all entries in MAF files as mutations present in the associated cell lines.

Merged mutation calls (coding region, germline filtered):
```{r, eval=FALSE}
maf_raw <- read.delim("./data_munging/18Q4/depmap_18Q4_mutation_calls_20181114.csv.gz", header = TRUE, sep = ",", row.names = 1)
colnames(maf_raw)[colnames(maf_raw) == "Tumor_Sample_Barcode"] <- "CCLE_Name"

# Select columns
maf_df <- subset(maf_raw, select = c("Hugo_Symbol", "CCLE_Name", "DepMap_ID", "Variant_Classification", "Reference_Allele", "Tumor_Seq_Allele1", "Genome_Change"))
maf_df$Reference_Allele_Length <- ifelse(maf_df$Reference_Allele == "-", 0, nchar(as.character(maf_df$Reference_Allele)))
maf_df$Tumor_Seq_Allele1_Length <- nchar(as.character(maf_df$Tumor_Seq_Allele1))

# Add Mutation_Status column
maf_df$Mutation_Status_Nonsilent <- ifelse(maf_df$Variant_Classification == "Silent", "Wiltype", "Mutant")
maf_df <- unique(maf_df)
saveRDS(maf_df, "./data_munging/rds/maf_df_18Q4.rds", compress = "xz")
```

```{r, eval=FALSE}
maf_df <- readRDS("./data_munging/rds/maf_df_18Q4.rds")
```

@@

DepMap WES CN Data:
```{r, eval=FALSE}
cn <- read.delim("./data_munging/18Q4/public_18Q4_gene_cn_20181115.csv.gz", sep = ",", check.names = FALSE, header = TRUE)

# Remove Entrez gene IDs from colnames
# Format: Hugo_Symbol (Entrez ID)
colnames(cn) <- gsub(" .*", "", colnames(cn))

# Melt
cn_melt <- melt(data = cn, id.vars = "DepMap_ID", measure.vars = colnames(cn[2:ncol(cn)]), variable.name = "Hugo_Symbol", value.name = "CN")

saveRDS(cn_melt, "./data_munging/rds/cn_melt_18Q4.rds", compress = "xz")
```

```{r, eval=FALSE}
cn_melt <- readRDS("./data_munging/rds/cn_melt_18Q4.rds")
```

@@

CCLE RNAseq gene expression data (log2(TPM+1)):
```{r, eval=FALSE}
ge <- read.delim("./../htscreen_lineages_giant_files/CCLE_depMap_18Q4_TPM_v2_20181115.csv.gz", header = TRUE, sep = ",", check.names = FALSE)

# Remove Entrez gene IDs from colnames
# Format: Hugo_Symbol (Entrez ID)
colnames(ge) <- gsub(" .*", "", colnames(ge))
colnames(ge)[1] <- "DepMap_ID"

# Melt
ge_melt <- melt(ge, id.vars = "DepMap_ID", measure.vars = colnames(ge)[2:ncol(ge)], variable.name = "Hugo_Symbol", value.name = "TPM")

saveRDS(ge_melt, "./data_munging/rds/ge_melt_18Q4.rds", compress = "xz")
```

```{r, eval=FALSE}
ge_melt <- readRDS("./data_munging/rds/ge_melt_18Q4.rds")
```

@@

```{r, eval=FALSE}
dep <- read.delim("./data_munging/18Q4/gene_dependency_18Q4_20181115.csv.gz", sep = ",", header  = TRUE, check.names = FALSE)
# Remove Entrez gene IDs from colnames
colnames(dep) <- gsub(" .*", "", colnames(dep))
```

@@

The latest CRISPR CERES score data (18Q3, August 2018) was pulled from the [DepMap Data Portal](https://depmap.org/portal/download/) (Broad Institute Cancer Dependency Map 2018, Meyers et al. 2017).
```{r, eval=FALSE}
crispr <- read.delim("./data_munging/18Q4/gene_effect_18Q4_20181115.csv.gz", sep = ",", header  = TRUE, check.names = FALSE)
# Remove Entrez gene IDs from colnames
colnames(crispr) <- gsub(" .*", "", colnames(crispr))
```

Merge annotation data:
```{r, eval=FALSE}
# Melt CRISPR dataset for merging
crispr_melt <- melt(crispr, id.vars = "DepMap_ID", measure.vars = colnames(crispr)[2:ncol(crispr)], variable.name = "Hugo_Symbol", value.name = "Score")
# Melt dependency probabilities dataset for merging
dep_melt <- melt(dep, id.vars = "DepMap_ID", measure.vars = colnames(dep)[2:ncol(dep)], variable.name = "Hugo_Symbol", value.name = "Dep_Prob")
# Merge dependency probabilities
crispr_melt <- merge(crispr_melt, dep_melt, by = c("DepMap_ID", "Hugo_Symbol"), all.x = TRUE)
# Merge cell line metadata
crispr_melt <- merge(crispr_melt, ccl_info, by = "DepMap_ID", all.x = TRUE)
crispr_melt <- merge(crispr_melt, crispr_meta, by = c("CCLE_Name", "DepMap_ID"), all.x = TRUE)
# Merge mutation annotations
maf_filt_crispr <- filter(maf_df, Hugo_Symbol %in% as.character(unique(crispr$Hugo_Symbol)))
crispr_muts <- merge(crispr_melt, maf_filt_crispr, by = c("Hugo_Symbol", "CCLE_Name", "DepMap_ID"), all.x = TRUE)
crispr_muts$Hugo_Symbol <- factor(crispr_muts$Hugo_Symbol)
crispr_muts <- crispr_muts %>% mutate(Mutation_Status_Nonsilent = if_else(is.na(Mutation_Status_Nonsilent), "Wildtype", Mutation_Status_Nonsilent))
# Summarize number of mutant and Wildtype cell lines
crispr_muts_summ <- crispr_muts %>% group_by(Hugo_Symbol) %>%
  summarize(N_Nonsilent_Wildtype = sum(Mutation_Status_Nonsilent == "Wildtype"),
            N_Nonsilent_Mutant = sum(Mutation_Status_Nonsilent == "Mutant"))
# Merge test results back into full dataset, which restores information lost in the summarization
crispr_data <- merge(crispr_muts_summ, crispr_muts, by = "Hugo_Symbol")
# Add Color columns
crispr_data$Color_Nonsilent <- ifelse(crispr_data$Mutation_Status_Nonsilent == "Wildtype", "thistle", "darkorchid")
crispr_data$Color_Nonsilent <- factor(crispr_data$Color_Nonsilent)
# Cell line lineages
crispr_data <- merge(crispr_data, ccl_converter, by = c("CCLE_Name", "DepMap_ID"), all.x = TRUE)
levels(crispr_data$lineage_name) <- sort(levels(crispr_data$lineage_name), decreasing = TRUE)
# Copy number
crispr_data <- merge(crispr_data, cn_melt, by = c("Hugo_Symbol", "DepMap_ID"), all.x = TRUE)
# Gene expression (RPKM)
ge_filt <- filter(ge_melt, Hugo_Symbol %in% as.character(unique(crispr_data$Hugo_Symbol)) & DepMap_ID %in% as.character(unique(crispr_data$DepMap_ID)))
crispr_data <- merge(crispr_data, ge_filt, by = c("Hugo_Symbol", "DepMap_ID", "CCLE_Name"), all.x = TRUE)
crispr_data$RPKM_log2 <- log2(crispr_data$RPKM + 0.0001)
saveRDS(crispr_data, "./../htscreen_lineages_giant_files/crispr_data_18Q3.rds", compress = "xz")
```

```{r}
crispr_data <- readRDS("./../htscreen_lineages_giant_files/crispr_data_18Q3.rds")

@@

```{r, eval=TRUE}
# Drug metadata
drug_meta <- read.delim("./data_munging/drug_id_drug_name_table_10_12_2018.tsv", sep = "\t", header = FALSE)
colnames(drug_meta) <- c("CID", "Drug")
drug_meta$Drug <- tolower(drug_meta$Drug)
maf_df_g2p <- filter(maf_df, Hugo_Symbol %in% g2p_genes)
ge_filt <- filter(ge_melt, Hugo_Symbol %in% g2p_genes)
cn_filt <- filter(cn_melt, Hugo_Symbol %in% g2p_genes)

# CCLE
ccle <- read.delim("./data_munging/data_drug_auc_ccle_10_12_2018.csv", sep = "\t", header  = TRUE, row.names = 1, check.names = FALSE)
ccle <- merge(drug_meta, ccle, by = "CID")
ccle_melt <- melt(ccle, id.vars = c("CID", "Drug"), measure.vars = colnames(ccle)[3:ncol(ccle)], variable.name = "accession_id", value.name = "AUC")
ccle_grid <- expand.grid("accession_id" = unique(ccle_melt$accession_id), "Hugo_Symbol" = g2p_genes)
ccle_data <- merge(ccle_grid, ccle_melt, by = "accession_id", all.x = TRUE)
ccle_data <- merge(ccle_data, ccl_converter, by = "accession_id", all.x = TRUE)
ccle_data <- merge(ccle_data, maf_df_g2p, by = c("Hugo_Symbol", "CCLE_Name", "DepMap_ID"), all.x = TRUE)
ccle_data$Mutation_Status_Nonsilent <- ifelse(is.na(ccle_data$Mutation_Status_Nonsilent), "Wildtype", ccle_data$Mutation_Status_Nonsilent)
ccle_data <- merge(ccle_data, ge_filt, by = c("Hugo_Symbol", "CCLE_Name", "DepMap_ID"), all.x = TRUE)
ccle_data <- merge(ccle_data, cn_filt, by = c("Hugo_Symbol", "DepMap_ID"), all.x = TRUE)
ccle_data$Hugo_Symbol <- factor(ccle_data$Hugo_Symbol)
ccle_data$AUC <- as.numeric(ccle_data$AUC)
ccle_data$Color_Nonsilent <- ifelse(ccle_data$Mutation_Status_Nonsilent == "Wildtype", "palegreen3", "springgreen4")
ccle_data$Color_Nonsilent <- factor(ccle_data$Color_Nonsilent)
ccle_data <- filter(ccle_data, !is.na(ccle_data$AUC))
ccle_data$Drug_Gene <- paste0(ccle_data$Drug, "_", ccle_data$Hugo_Symbol)
saveRDS(ccle_data, "./data_munging/rds/ccle_data_18Q3_g2pgenes.rds", compress = "xz")

# CTRP
ctrp <- read.delim("./data_munging/data_drug_auc_ctrp_10_12_2018.csv", sep = "\t", header  = TRUE, row.names = 1, check.names = FALSE)
ctrp <- merge(drug_meta, ctrp, by = "CID")
ctrp_melt <- melt(ctrp, id.vars = c("CID", "Drug"), measure.vars = colnames(ctrp)[3:ncol(ctrp)], variable.name = "accession_id", value.name = "AUC")
ctrp_grid <- expand.grid("accession_id" = unique(ctrp_melt$accession_id), "Hugo_Symbol" = g2p_genes)
ctrp_data <- merge(ctrp_grid, ctrp_melt, by = "accession_id", all.x = TRUE)
ctrp_data <- merge(ctrp_data, ccl_converter, by = "accession_id", all.x = TRUE)
ctrp_data <- merge(ctrp_data, maf_df_g2p, by = c("Hugo_Symbol", "CCLE_Name", "DepMap_ID"), all.x = TRUE)
ctrp_data$Mutation_Status_Nonsilent <- ifelse(is.na(ctrp_data$Mutation_Status_Nonsilent), "Wildtype", ctrp_data$Mutation_Status_Nonsilent)
ctrp_data <- merge(ctrp_data, ge_filt, by = c("Hugo_Symbol", "CCLE_Name", "DepMap_ID"), all.x = TRUE)
ctrp_data <- merge(ctrp_data, cn_filt, by = c("Hugo_Symbol", "DepMap_ID"), all.x = TRUE)
ctrp_data$Hugo_Symbol <- factor(ctrp_data$Hugo_Symbol)
ctrp_data$AUC <- as.numeric(ctrp_data$AUC)
ctrp_data$Color_Nonsilent <- ifelse(ctrp_data$Mutation_Status_Nonsilent == "Wildtype", "slategray3", "steelblue4")
ctrp_data$Color_Nonsilent <- factor(ctrp_data$Color_Nonsilent)
ctrp_data <- filter(ctrp_data, !is.na(ctrp_data$AUC))
ctrp_data$Drug_Gene <- paste0(ctrp_data$Drug, "_", ctrp_data$Hugo_Symbol)
saveRDS(ctrp_data, "./data_munging/rds/ctrp_data_18Q3_g2pgenes.rds", compress = "xz")
